Simulations of gravity-driven flow of binary liquids in microchannels

### 2011-03-22 Text/paper.tex
Continuing with the revisions.

The procedure can be described as follows:
\begin{description}
 \item[Capillary number choice] Blah-blah
 \item[Film thickness] You choose film thickness. 
 \item[Grid choice]
 \item[Body force] Blah-blah
\end{description}

The film thickness for such small capillary numbers \cite{bretherton} is
given by the equation:
  \begin{equation}
  \frac{h_{\infty}}{H_{\mathrm{eff}}}=0.6687 \Ca^{2/3}=0.0194.
  \end{equation}
Thus, the film
 occupies about $2$ percent of the whole channel height. Let us impose that $8$ lattice
units are required to resolve  $2$ percent of the channel. Note that $5$ out of $8$ these lattice
units
($62.5$ percent) are used to resolve the phase interface. Therefore the channel width is $400$
lattice units.

 The film thickness for such capillary number is taken from Fig.~\ref{fig:heil:planar},
  and for Reynolds number $40$ equals $0.12$. We want to
  resolve $6\%$ of the whole channel width with $12$ lattice units, where $5$
lattice units are related to the interface resolution.  Therefore, the channel
  width is $200$ lattice units and the grid size is $200 \times 3000$.

We take the bubble length equal to $5$ channel widths
  and the distance between bubbles as $3$ bubble lengths. This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with an interface thickness of $62.5$ percent  of the
film thickness, one gets (in 2D) a memory
requirement of $345 \mathrm{MB}$.

After resolving the interface one can  calculate the
  velocity from the capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.005 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.005
\frac{0.0377}{0.6666}=2.82 \cdot10^{-4},
  \end{equation}
  where $\mu_{\mathrm{liq}}=1/3 (2.5-0.5)=0.6666$.
  This velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
  \begin{equation}
  \begin{aligned}
  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
  \end{aligned}
  \end{equation}
Note that the velocity is proportional to the physical time step (small velocities imply longer
running
simulations):
\begin{equation}
\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
\end{equation}
where, for example, the real velocities in the physical world can reach
approximately $1$ m/s and the real height up to $1$ mm. In this case one
can easily approximate the physical time step during one lattice Boltzmann iteration:
\begin{equation}
\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
\end{equation}
which is small. A typical simulation runs approximately
$100000-300000$ time steps due to reach steady-state. That gives 
physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
probably not enough for the physical system to come to the pseudo-steady state.

@@

  The velocity can be
calculated from the capillary number as:
  \begin{equation}
  U_{\mathrm{bubble}}=0.05 \frac{\gamma}{\mu_{\mathrm{liq}}}=0.05 \frac{0.037712}{0.6666}=2.82\cdot
  10^{-3}.
  \end{equation}
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}

@@

\subsection{Grid refinement}
To properly estimate the interface resolution one needs to study the convergence as a function of
the grid resolution. To do that the grid resolution is varied while all remaining parameters,
including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
ratio of the interface thickness to the 
film thickness at which results are no longer dependent on the grid resolution.

>>>>

Overall, the setting simulations procedure can be described in steps as follows:
\begin{description}
 \item[Capillary number] One first needs to be supplied with the capillary number for
simulations. For the demonstration purposes we chose $\Ca=0.05$. 
 \item[Film thickness] After the capillary number is prescribed, one needs to approximate the film
thickness $\delta$. It can be done either by taking correlations from numerical simulations by
\citet{heil-bretherton} and \citet{giavedoni-numerical} or by using the classical Bretherton
correlation for small cappillary numbers $\delta=0.6687 \Ca^{2/3}$. In the case of $\Ca=0.05$, the 
film thickness is taken from {\color{green} Fig.~\ref{fig:heil:planar}},
  and for Reynolds number $40$ equals $0.06$.  
\item[Grid choice] After pecifying the film thickness, one needs to choose the associate
number of nodes to resolve the film thickness. Note that for paremeters indicated above the
interface thickness is spread over approximately $5$ lattice units. The study of the interface
resolution to the film thickness, Section \ref{section:grid:refinement}, suggests to choose the
film thickness to be $2-2.5$ times larger than the interface thickness. Therefore, we want to
  resolve $6\%$ of the whole channel width with $12$ lattice units.  Thus, the channel
  width is $200$ lattice units. We take the bubble length equal to $5$ channel widths for the film
thickness to establish
  and the distance between bubbles as $3$ bubble lengths to avoid mutual influence of bubbles on
each other. Thus, the whole grid size is $200 \times 3000$.
 \item[Velocity] The relaxation parameters are taken to be in the stable range for simulations.
In the present simulations the viscosity ratio is $\dfrac{\mu_{liq}}{\mu_{gas}}=10$ with the
correspoding relaxation times $\tau_{liq}=2.5$ and $\tau_{gas}=0.7$. Given the viscosity of the
liquid one can obtain the velocity of the bubble from the capillary number: 
\begin{equation}
\begin{aligned}
&U_{\mathrm{bubble}}=\Ca \frac{\gamma}{\mu_{\mathrm{liq}}}=\Ca \frac{\sqrt{\frac{8 k
A}{9}}}{\frac{1}{3}(\tau_{\mathrm{liq}}-\frac{1}{2})}\\
&U_{\mathrm{bubble}}=0.05 \frac{0.0377}{0.6666}=2.82 \cdot10^{-3},
\end{aligned}
\end{equation}
where $\mu_{\mathrm{liq}}=1/3 (2.5-0.5)=0.6666$.
\item[Body force] 
  The velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution:
  \begin{equation}
  \begin{aligned}
  &\frac{\mathrm{d}P}{\mathrm{d}x}=\frac{8}{N_y^2}\gamma \Ca=3.77\cdot10^{-7}.
  \end{aligned}
  \end{equation}

After resolving the interface one can  calculate the
  velocity from the capillary number as:
  This velocity implies the following pressure gradient calculated from the
  Poiseuille flow solution, Eq. \ref{poiseuille:velocity:center}: 
  \begin{equation}
  \begin{aligned}
  \frac{\mathrm{d}P}{\mathrm{d}x}=9.42 \cdot 10^{-9}.
  \end{aligned}
  \end{equation}
Note that the velocity is proportional to the physical time step (small velocities imply longer
running
simulations):
\begin{equation}
\Delta t =\frac{U_{\mathrm{LBM}}}{U_{\mathrm{phys}}} \Delta x ,
\end{equation}
where, for example, the real velocities in the physical world can reach
approximately $1$ m/s and the real height up to $1$ mm. In this case one
can easily approximate the physical time step during one lattice Boltzmann iteration:
\begin{equation}
\Delta t = 7.05 \cdot 10^{-10} \,\mathrm{s}
\end{equation}
which is small. A typical simulation runs approximately
$100000-300000$ time steps due to reach steady-state. That gives 
physical time simulation around $10^{-5}-10^{-4}\,\mathrm{s}$, which is
probably not enough for the physical system to come to the pseudo-steady state.

\end{description}

This gives a
grid size of $400 \times 6000$. In terms of computer memory
requirements, the calculations with double precision accuracy demand:
\begin{equation}
\begin{aligned}
&2\text{ sets of populations } \times \,\, 8\,\, \text{bytes per population
} \times \,\, 9 \text{
populations}\\
&\times\,400\,\times\,6000\approx 345 \mathrm{MB}. 
\end{aligned}
\end{equation}
Even with an interface thickness of $62.5$ percent  of the
film thickness, one gets (in 2D) a memory
requirement of $345 \mathrm{MB}$.

@@

@@

\subsection{Grid refinement}
\label{section:grid:refinement}
To properly estimate the interface resolution one needs to study the convergence as a function of
the grid resolution. To do that the grid resolution is varied while all remaining parameters,
including the bubble velocity and the capillary number, are fixed.  Our goal is to determine the 
ratio of the interface thickness to the 
film thickness at which results are no longer dependent on the grid resolution.
